---
title: "Homework 5"
author: "Deanna Springgay"
date: "4/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(forecast)
library(TSA)
library(tseries)
set.seed(64584)
```

### Problems

Written questions are attached at the end

### 8.6

a) The residuals seem to be random around a zero mean.
```{r 8.6a, echo=FALSE}
series1=arima.sim(n=48,list(ar=c(1.5,-0.75)))
model1=arima(series1,order=c(2,0,0))
plot(rstandard(model1), type='o'); abline(h=0)
```

b) The residuals seem to follow a normal distribution.
```{r 8.6b, echo=FALSE}
qqnorm(rstandard(model1))
qqline(rstandard(model1))
```

c) None of the lags are significant.
```{r 8.6c, echo=FALSE}
acf(rstandard(model1))
```

d) From the Ljung-Box test we fail to reject the null hypothesis of independent residuals.
```{r 8.6d, echo=FALSE}
LB.test(model1,lag=12)
```
 
### 8.9

While both models are significant, the arima(0,1,1) model has a better log likelihood and AIC.
```{r 8.9a, echo=FALSE}
data(robot)
model2=arima(robot,order=c(1,0,0))
residual2=rstandard(model2)
model3=arima(robot,order=c(0,1,1))
residual3=rstandard(model3)
model2
model3
```

The arima(1,0,0) model might have some drift, but arima(0,1,1) doesn't seem to have any problems at this point.
```{r 8.9b, echo=FALSE}
plot(residual2,ylab='arima(1,0,0)'); abline(h=0)
plot(residual3,ylab='arima(0,1,1)'); abline(h=0)
```

The arima(1,0,0) model has a lag problem with many lags being significant. The arima(0,1,1) model doesn't have any significant lags except for lag100 and the Ljung-Box test gives a good result of failing to reject the null hypothesis of not showing lack of fit. We will only move forward with the arima(0,1,1) model from this point on.
```{r 8.9c, echo=FALSE}
acf(residuals(model2), main='arima(1,0,0)',ylab='ACF'); LB.test(model2)
acf(residuals(model3), main='arima(0,1,1)',ylab='ACF'); LB.test(model3)
```

From the following outputs, we can conclude that the arima(0,1,1) model fits well - normality is good, the residuals are also good, and the p-values are acceptable.
```{r 8.9d, echo=FALSE}
qqnorm(residuals(model3)); qqline(residuals(model3))
shapiro.test(residuals(model3))
tsdiag(model3)
```

### 9.16

a)
```{r 9.6a, echo=FALSE}
series2=(arima.sim(n=45,list(order=c(0,2,2),ma=c(-1,0.75)))[-1])[-1]
actual=window(series2,start=41); series2=window(series2,end=40)
model4=arima(series2,order=c(0,2,2)); model4
```

b) The forcasts are linear when the rest of the plot isn't.
```{r 9.6b, echo=FALSE}
result=plot(model4,n.ahead=5,col=NULL)
```

c) The forecast appears to be underestimating the actual values.
```{r 9.6c, echo=FALSE}
forecast=result$pred; cbind(actual,forecast)
```

d) The actual values fall within the 95% CI except the point at 42 which appears to be on the CI boundary. From the actual values, time 42 is just below the upper CI bound.
```{r 9.6d, echo=FALSE}
plot(model4,n1=38,n.ahead=5)
points(x=seq(41,45),y=actual)
lower=result$lpi; upper=result$upi; cbind(lower,actual,upper)
```

e)
```{r 9.6e, echo=FALSE}
# series3=(arima.sim(n=45,list(order=c(1,1,2),ma=c(-1,0.75)))[-1])[-1]
# actual2=window(series3,start=41); series3=window(series3,end=40)
# model5=arima(series3,order=c(1,1,2)); model5
# 
# result2=plot(model5,n.ahead=5,col=NULL)
# 
# forecast2=result2$pred; cbind(actual2,forecast)
# 
# plot(model5,n1=38,n.ahead=5)
# points(x=seq(41,45),y=actual2)
# lower2=result2$lpi; upper2=result2$upi; cbind(lower2,actual2,upper2)
```

### 9.23

a)
```{r 9.23a, echo=FALSE}
data(robot)
model6=arima(robot,order=c(0,1,1))
result3=plot(model6,n.ahead=5,col=NULL)
predicted=result3$pred
upper=result3$upi
lower=result3$lpi
cbind(lower,predicted,upper)
```

b) The forcasts are basically constant and the CI is very wide, likely due to the variance seen in previous values.
```{r 9.23b, echo=FALSE}
plot(model6,n1=300,n.ahead=5)
```

c) Both of these models produce very similar forecasts.
```{r 9.23c, echo=FALSE}
model7=arima(robot,order=c(1,0,1))
result4=plot(model7,n.ahead=5,col=NULL)
predicted=result4$pred
upper=result4$upi
lower=result4$lpi
cbind(lower,predicted,upper)
plot(model7,n1=300,n.ahead=5)
```