---
title: "Hw 1 R Code"
author: "Deanna Springgay"
date: "1/31/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Libraries
library(tidyverse)
library(TSA)
```

### 1.4

Below are two plots of different simulations of the chi-squared distribution in following code block. Both appear to follow non-normal distributions but they are not random - they follow a chi-squared distribution.

```{r chi-square}
set.seed(7)
chiSquare <- ts(rchisq(n = 48, df = 2))
plot(chiSquare, type = "o", ylab = "", main = "Chi-Squared Distribution, n = 48, df = 2")
abline(h = 0, lty = 2)
```

```{r chi-square_1, echo=FALSE}
set.seed(3)
chiSquare <- ts(rchisq(n = 48, df = 2))
plot(chiSquare, type = "o", ylab = "", main = "Chi-Squared Distribution, n = 48, df = 2")
abline(h = 0, lty = 2)
```

### 1.5

Below are two plots of different simulations of the t distribution in following code block. Both appear to follow normal and do not appear to be random.

```{r t-dist}
set.seed(17)
chiSquare <- ts(rt(n = 48, df = 5))
plot(chiSquare, type = "o", ylab = "", main = "t Distribution, n = 48, df = 5")
abline(h = 0, lty = 2)
```

```{r t-dist_1, echo=FALSE}
set.seed(90)
chiSquare <- ts(rt(n = 48, df = 5))
plot(chiSquare, type = "o", ylab = "", main = "t Distribution, n = 48, df = 5")
abline(h = 0, lty = 2)
```

### 1.6

```{r dubuque}
data("tempdub")
plot(tempdub, type = "l", ylab = "Average Temperature")
points(y = tempdub, x = time(tempdub), pch = as.vector(season(tempdub)))
```